---
title: "hw05"
author: "Matthew Clagett"
date: "April 27, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

auxillary function for die
```{r}
#' @title check_sides
#' @description checks the number of sides of a die to see if it is a value argument
#' @param input sides of the die
#' @return 
check_sides <- function(sides){
     if(length(sides) != 6 ){
          stop("sides must be a vector length of 6")
     }
}

#' @title check_prob
#' @description checks the probabilities to see if they are a correct input
#' @param input probabilities of the die sides
#' @return 
check_prob <- function(prob){
     if(sum(prob) != 1){
          stop("elements in prob must add up to 1")
     } else if(length(prob) != 6){
          stop("prob must be a vector length of 6")
     }
}
```
print method for die
```{r}
#' @title print.die
#' @description print method for objects of type "die"
print.die <- function(x, ...){
     cat('object "die"\n\n')
     print(data.frame(side = x$side, prob = x$prob))
}
```

Created function that builds a die object
```{r}
#' @title die object
#' @description creates an object of class "die"
#' @param input sides of each die - six total
#' @param input probabilities corresponding to each side of the die
die <- function(sides = c(1:6), prob = rep(1/length(sides), length(sides))){
     check_sides(sides)
     check_prob(prob)
     
     die1 <- list(side = sides,
                  prob = prob)
     die1 <- structure(die1, class = "die")
     die1
}
```


auxillary function for rolls
```{r}
#' @title check_times
#' @description checks whether input is a valid number of rolls for a die
#' @param times number of times the die is to be rolled
check_times <- function(times){
     if(times < 1 | times%%1!=0){
          stop("die roll must be a positive integer")
     }
}
```

print method for roll
```{r}
#' @title print.roll
#' @description print method for objects of type roll
print.roll <- function(x, ...){
     cat('object "roll"\n\n')
     print(x$rolls)
}
```
function creates an object class roll
```{r}
#' @title roll
#' @description takes a die and a number of rolls, returns object class rolls
#' @param die object of die class
#' @param times number of rolls
roll <- function(die, times){
     check_times(times)
     rolls <- list(rolls = sample(die$side, times, replace=TRUE, prob = die$prob),
                   sides = die$side,
                   prob = die$prob,
                   total = times)
     rolls <- structure(rolls, class = "roll")
     rolls
}
```
testing the functionality of roll

summary method for roll
```{r}
#' @title summary.roll
#' @description summary method for objects of type roll
summary.roll <- function(x, ...){
     summary <- list(freqs = data.frame(count = table(x$rolls),
                prop = x$prob))
     colnames(summary$freqs) <- c("side", "count", "prop")
     structure(summary, class="summary.roll")
}
```
print summary method for roll
```{r}
#' @title print.summary.roll
#' @description print method for summary.roll
print.summary.roll <- function(x, ...){
     cat('summary "roll"\n\n')
     print(x$freqs)
}
```

plot method for roll
```{r}
plot.roll <- function(x, ...){
     relative_freqs <- x$freqs$count/sum(x$freqs$count)
     barplot(table(x$rolls)/length(x$rolls), 
             xlab = "sides of die",
             ylab = "relative frequencies", 
             main = paste0("Frequencies in a series of ",length(x$rolls)," rolls"))
}
```

extraction method for roll
```{r}
#' @title extraction.roll
#' @description extraction method for object of class "roll"
"[.roll" <- function(x, i){
     x$rolls[i]
}
```


replacement method for roll
```{r}
#' @title index roll
#' @description index method for summary.roll
"[<-.roll" <- function(x,i,value){
     if(!(value %in% x$sides)){
          stop(paste('\nReplacing value must be in: ',x$sides))
     }
     if(i>x$total){
          stop("\nindex out of bounds")
     }
     x$rolls[i] <- value
     x
}
```

```{r}
#' @title +.roll
#' @description addition method for roll
"+.roll" <- function(obj, incr){
     if(length(incr) != 1 | incr <= 0){
          stop("\ninvalid addition (must be positive")
     }
     more_rolls <- roll(obj, times = incr)
     obj$rolls <- c(obj$rolls, more_rolls$rolls)
     obj$total <- obj$total + incr
     obj
}
```

### Testing functionality of the "dieroller" package
Problem 1:
1000 games of rolling a die 4 times, find the probability of at least one six

generating a fair die
```{r}
die <- die()
```

Creating a simulation of 100 rolls
```{r}
rolls <- matrix(nrow=1000, ncol =4)
for (i in 1:1000){
     rolls[i,] <- roll(die, 4)$rolls
}
```
finding the relative frequencies of finding a 6
```{r}
rf <- apply(rolls, 1, function(x){6 %in% x})

relative_freqs <- rep(0,1000)
for (i in 1:1000){
     relative_freqs[i] <- sum(rf[1:i]/length(rf[1:i]))
}
relative_freqs
```

roll two dice, 24 times find frequencies of both having sixes
```{r}
sixes <- c(rep(0,1000))
for (i in 1:1000){
     sixes[i] <- any(roll(die,24)[1:24] == 6 & roll(die,24)[1:24] == 6)
}


relative_freqs_both <- rep(0,1000)
for (i in 1:1000){
     relative_freqs_both[i] <- sum(sixes[1:i]/length(sixes[1:i]))
}
relative_freqs_both
```


